---
title: "Figure 3c"
author: "Johannes S. P. Doehl"
date: "2024-07-02"
output: pdf_document
---

~~~{=comment}
# Setup
~~~
```{r setup-figure-3c, echo = FALSE, include = FALSE, eval = TRUE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE, eval = TRUE, cache = TRUE, dev = "png", warning = FALSE, message = FALSE, fig.pos = "H", out.extra = "")
options(java.parameters = "-Xmx200000m")
Sys.setenv(JAVA_HOME='C:\\Program Files\\Java\\jre-1.8') # Needed to get around loading error of rJava
```

```{r extract-relevant-data-for-figure-3c}
# Figure name
FigNam <- "Figure 3c"

# Subset data that applies to this figure
SubDat <- ResSur %>%
  .[names(.) %like% FigNam] %>%
  .[[1]]

# Total N
TotN <- DatSur %>%
  .[names(.) %like% FigNam] %>%
  .[[1]] %>%
  nrow(.)

# Group N
GrpN <- table(DatSur %>%
  .[names(.) %like% FigNam] %>%
  .[[1]] %>%
  .$Diet_Route) %>%
  paste(names(.), ., sep = "=", collapse = ", ")
```

Here, we analyzed the occurrence of ocular pathology following *Leishmania donovani* infection in well-nourished (WN) and malnourished (MN) BALB/c mice. A total of *N*=`r TotN` BALB/c mice (`r GrpN`) were examined on a weekly bases post infection by either "needle" or sand fly (SF) route and occurrence of pathology was recorded as time-to-event data. 

#### Survival analysis

The data was analyzed by the Mantel-Haenszel's log-rank test by use of the survdiff() function from the survival package in R. The test output is shown in table `r TabNum + 1`, which was statistically significant (F(`r length(SubDat[["Survival"]][["obs"]]) - 1`)=`r round(SubDat[["Survival"]][["chisq"]], 2)`, `r if (SubDat[["Survival"]][["pvalue"]] < 0.0001) { "<0.0001" } else { format(round(SubDat[["Survival"]][["pvalue"]], 4), scientific = FALSE) }`).

```{r table-99-mantel-haenszel-s-log-rank-test, include = TRUE, eval = TRUE}
# Mantel-Haenszel log-rank test
cbind("N" = SubDat[["Survival"]][["n"]], 
      "Observed" = SubDat[["Survival"]][["obs"]],
      "Expected" = SubDat[["Survival"]][["exp"]],
      "(O-E)^2/E" = (SubDat[["Survival"]][["obs"]] - 
                     SubDat[["Survival"]][["exp"]])^2 / 
                     SubDat[["Survival"]][["exp"]],
      "(O-E)^2/V" = diagonals((SubDat[["Survival"]][["obs"]] - 
                               SubDat[["Survival"]][["exp"]])^2 / 
                               SubDat[["Survival"]][["var"]])) %>%
  as.data.frame(.) %>%
  mutate(., across(where(is.numeric), ~ round(.x, 4))) %>%
  set_rownames(str_split(rownames(.), ", ") %>%
                 lapply(., function(zz) {
                   gsub(".*=", "", zz) %>%
                     paste(., collapse = "_") %>%
                     str_trim(., "right")
                   }) %>%
                 unlist(.)) %>%
  # Build a table
  gt(rownames_to_stub = TRUE) %>%
    tab_stubhead(label = "Groups") %>%
    tab_header(
      title = md("**Appendix Table 99**"),
      subtitle = md("**Log-rank test**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    ) %>%
    tab_spanner(
      label = "Chi-Square",
      columns = 5
    ) %>%
    tab_spanner(
      label = "log-rank",
      columns = 6
    )
```

The pairwise Mantel-Haenszel' log-rank test, which was adjusted by the Benjamin-Hochberg correction, showed that the malnourished, sand fly inoculated group was statistically significant from all other groups, while there was no statistically significant difference between the remaining three groups (Table `r TabNum + 2`). 

```{r table-100-pairwise-mantel-haenszel-s-log-rank-test, include = TRUE, eval = TRUE}
# Pairwise log-rank test
SubDat[["Pairwise"]][["p.value"]] %>%
  as.data.frame(.) %>% 
  mutate(., across(where(is.numeric), ~ round(.x, 4))) %>%
  mutate(., across(where(is.numeric), ~ ifelse(.x < 0.0001, "<0.0001", format(.x, scientific = FALSE)))) %>%
  set_rownames(str_split(rownames(.), ", ") %>%
                 lapply(., function(zz) {
                   gsub(".*=", "", zz) %>%
                     paste(., collapse = "_") %>%
                     str_trim(., "right")
                   }) %>%
                 unlist(.)) %>%
  set_colnames(str_split(colnames(.), ", ") %>%
                 lapply(., function(zz) {
                   gsub(".*=", "", zz) %>%
                     paste(., collapse = "_") %>%
                     str_trim(., "right")
                   }) %>%
                 unlist(.)) %>%
  # Build a table
  gt(rownames_to_stub = TRUE) %>%
    tab_stubhead(label = "Groups") %>%
    tab_header(
      title = md("**Appendix Table 100**"),
      subtitle = md("**Pairwise Log-rank test**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    )
```

In fact, only malnourished animals developed ocular pathology post infection and the grand majority of these were inoculated by sand fly (Table `r TabNum + 3`), suggesting that sand flies significantly increase the occurrence of ocular pathology in malnourished hosts.

```{r table-101-percent-ocular-pathology-rate, include = TRUE, eval = TRUE}
cbind(
  cbind(
    "Group N" = table(DatSur %>%
      .[names(.) %like% FigNam] %>%
      .[[1]] %>%
      .$Diet_Route),
      
    "Ocular Pathology" = table(DatSur %>%
      .[names(.) %like% FigNam] %>%
      .[[1]] %>%
      .$Event, 
      DatSur %>%
      .[names(.) %like% FigNam] %>%
      .[[1]] %>%
      .$Diet_Route) %>%
      .["1",]),
  
  table(DatSur %>%
    .[names(.) %like% FigNam] %>%
    .[[1]] %>%
    .$Event, 
    DatSur %>%
    .[names(.) %like% FigNam] %>%
    .[[1]] %>%
    .$Diet_Route) %>%
    { .["1",] * 100 / table(DatSur %>%
    .[names(.) %like% FigNam] %>%
    .[[1]] %>%
    .$Diet_Route) } %>%
    as.data.frame.model.matrix(.) %>%
    mutate(., across(where(is.numeric), ~ format(round(.x, 2), nsmall = 2))) %>%
    mutate(., across(".", ~ paste0(.x, "%"))) %>%
    set_colnames("Occurrance Rate") %>%
    tibble::remove_rownames(.)
) %>%
  t(.) %>%
  as.data.frame(.) %>%
  .[, c(1,4,2,3)] %>%
  # Build a table
  gt(rownames_to_stub = TRUE) %>%
    tab_header(
      title = md("**Appendix Table 101**"),
      subtitle = md("**Ocular pathology occurrence rate**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    ) %>%
    cols_label(
      ends_with("Needle") ~ "Needle",
      ends_with("SF") ~ "Sand fly"
    ) %>%
    tab_spanner(
      label = "Well-nourished",
      columns = c(2,3)
    ) %>%
    tab_spanner(
      label = "Malnourished",
      columns = c(4,5)
    )
```

#### Cox proportional hazards regression model

```{r cox-proportional-hazards-regression-model-data}
# Subset data of Cox proportional hazards regression model that applies to this figure
SubDat <- ResCox %>%
  .[names(.) %like% FigNam] %>%
  .[[1]]
```

We also explored the data by Cox proportional hazards regression. Due to the lack of events in the well-nourished group, we had to resort to Firth's penalized maximum likelihood bias reduction method for Cox regression. The output showed that both predictors, "Diet" and "Route", were statistically significant (Table `r TabNum + 4`).

```{r table-102-Cox-PH-model-output, include = TRUE, eval = TRUE}
# Firth's penalized maximum likelihood bias reduction method for Cox regression 
summary(ResCox[["Figure 3c KM"]][["Firth's Cox regression (2 var)"]]) %>% 
  capture.output(.) %>% # make sure your window is wide enough
  .[c(7:9)] %>%
  trimws(.) %>%
  strsplit(., " ") %>%
  lapply(., function(yy) {
      
      yy[yy != ""]
  }) %>%
  lapply(., function(zz) {
      
      aa <- zz %>%
          as.data.frame(.) %>%
          t(.) %>%
          as.data.frame(.)
      
      if (length(aa) > 8) {
        
        aa %>%
          unite(., V5, c(V4, V5), sep = " ") %>%
          unite(., col = V7, c(V6, V7), sep = " ") %>%
          mutate(., V0 = ".", .before = "V1") %>%
          set_colnames(paste0("V", seq(1,8)))
        
       } else {
         aa
       }
  }) %>%
  bind_rows(.) %>%
  janitor::row_to_names(row_number = 1) %>%
  tibble::remove_rownames(.) %>%
  tibble::column_to_rownames(., var = ".") %>%
  mutate(., across(where(Hmisc::all.is.numeric), as.numeric)) %>%
  mutate(., "sig." = .[["p"]] %>%
           symnum(.,
                  cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 0.1, 1),
                  symbols = c("****", "***", "**", "*", "+", "ns"))) %>%
  mutate(., across("p", ~ ifelse(.x < 0.0001, "<0.0001", format(round(.x, 4), scientific = FALSE)))) %>%
  mutate(., across(where(is.numeric), ~ round(.x, 2))) %>%
  # Build a table
  gt(rownames_to_stub = TRUE) %>%
    tab_stubhead(label = "Groups") %>%
    tab_header(
      title = md("**Appendix Table 102**"),
      subtitle = md("**Firth's penalized maximum likelihood bias reduction method for Cox regression**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    )
```

The odds ratios suggested that "Diet" was a much more potent predictor for ocular pathology, than "Route", although both were statistically significant. Malnourished mice were about `r round(ResCox[["Figure 3c KM"]][["Hazard Ratio (2 var)"]]["DietMN", "AHR"],1)`-times more likely to develop ocular pathology than well-nourished mice compared to sand fly inoculated mice being `r round(ResCox[["Figure 3c KM"]][["Hazard Ratio (2 var)"]]["RouteSF", "AHR"],1)`-times more likely to developed ocular pathology compared to the once inoculated by needle (Table `r TabNum + 5`).

```{r table-103-odds-ratios, include = TRUE, eval = TRUE}
# Odds ratios
ResCox[["Figure 3c KM"]][["Hazard Ratio (2 var)"]] %>%
  mutate(., "sig." = .[["p-value"]] %>%
           symnum(.,
                  cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 0.1, 1),
                  symbols = c("****", "***", "**", "*", "+", "ns"))) %>%
  mutate(., across("p-value", ~ ifelse(.x < 0.0001, "<0.0001", format(round(.x, 4), scientific = FALSE)))) %>%
  mutate(., across(where(is.numeric), ~ round(.x, 2))) %>%
  # Build a table
  gt(rownames_to_stub = TRUE) %>%
    tab_stubhead(label = "Groups") %>%
    tab_header(
      title = md("**Appendix Table 103**"),
      subtitle = md("**Odds ratios**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    )
```

The pairwise comparison clearly confirmed the observation of the log-rank test that the malnourished, sand fly inoculated group was statistically significantly different from the other three groups (Table `r TabNum + 6` and `r TabNum + 7`). Here, the analysis also separated well-nourished needle inoculated mice from the other three groups, which may be due to the small samples.

```{r table-104-pairwise-comaprison, include = TRUE, eval = TRUE}
# Pariwise comparison by estimated marginal means
ResCox[["Figure 3c KM"]][["Pairwise comparison (2 var)"]][["Pairwise comparison"]] %>%
  as.data.frame(.) %>%
  mutate(., "sig." = .[["p.value"]] %>%
           symnum(.,
                  cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 0.1, 1),
                  symbols = c("****", "***", "**", "*", "+", "ns"))) %>%
  mutate(., across("p.value", ~ ifelse(.x < 0.0001, "<0.0001", format(round(.x, 4), scientific = FALSE)))) %>%
  mutate(., across(where(is.numeric), ~ round(.x, 2))) %>%
  # Build a table
  gt(rowname_col = "contrast") %>%
    tab_stubhead(label = "Groups") %>%
    tab_header(
      title = md("**Appendix Table 104**"),
      subtitle = md("**Pariwise comparison by estimated marginal means**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    )
```

```{r table-105-pairwise-comaprison-letters, include = TRUE, eval = TRUE}
ResCox[["Figure 3c KM"]][["Pairwise comparison (2 var)"]][["Pairwise letter code"]] %>%
  mutate(., across(where(is.numeric), ~ round(.x, 2))) %>%
  # Build a table
  gt(rowname_col = "contrast") %>%
    tab_stubhead(label = "Groups") %>%
    tab_header(
      title = md("**Appendix Table 105**"),
      subtitle = md("**Pariwise comparison by estimated marginal means (letters)**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    )
```


```{r update-table-number-figure-3c}
TabNum <- TabNum + 7
```
