---
title: "Figure S1a"
author: "Johannes S. P. Doehl"
date: "2024-06-07"
output: pdf_document
---

~~~{=comment}
# Setup
~~~
```{r setup-figure-S1a, echo = FALSE, include = FALSE, eval = TRUE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE, eval = TRUE, cache = TRUE, dev = "png", warning = FALSE, message = FALSE, fig.pos = "H", out.extra = "")
options(java.parameters = "-Xmx200000m")
Sys.setenv(JAVA_HOME='C:\\Program Files\\Java\\jre-1.8') # Needed to get around loading error of rJava
```

```{r extract-relevant-data-for-figure-S1a}
# Figure name
FigNam <- "Figure S1a"

# Subset data that applies to this figure
SubDat <- Res2wM %>%
  .[names(.) %like% FigNam] %>%
  purrr::set_names(FigNam)

# Total N per experiment
TotN <- lapply(Dat2wMSpt, function(yy) {
  
  nlevels(yy$Mouse)
  
})

# Group N
GrpN <- lapply(Dat2wMSpt, function(qq) {
  
  eval(parse(text = paste0("table(", paste(paste0("qq$", c("Diet", "Route", "Time_point")), collapse = ","), ")[,,1]")))
  
}) %>%
  unlist(.) %>%
  paste(names(.) %>%
          gsub(".*2.", "", .),
        ., 
        sep = "=",
        collapse = ", ")
```

## Data analysis

In figure S1a, we present the longitudinal weekly observation of mouse body weight pre *Leishmania donovani* infection, to assess the impact of well-nourishing and malnourishing diets on the BALB/c mice. Information of a total of *N*=`r TotN` BALB/c mice (`r GrpN`) over the course of 7 weeks are shown here; "Week_0" being the weight before shipment, "Week_7" being the final weight before infection.

We needed to analyze the data with a two-way approach to account for the two predictors, "Time_point" was the within-subject factor, while "Diet" was the between-subject factors in the analysis with "Weight_g" being the dependent outcome variable.

For a two-way mixed ANOVA, we had to assess the data for compliance with assumptions:

-   Data normality
-   Homogeneity of variance
-   Homogeneity of Covariance
-   No significant outliers
-   Assumption of sphericity


### Assumption analyses

#### Data normality

The assessment of the untransformed data distribution for each group was conducted by Shapiro-Wilks test and QQ-plot after splitting the data by both predictors. Note that all groups consisted of *N*=`r GrpN` individuals, which made groups large enough to assess data distribution reliably by Shapiro-Wilks test. Thus, we performed the analyses by Shapiro-Wilks test (Appendix table `r TabNum + 1`) and QQ-pots (Fig.S1a-1) and found deviations from normality only at the Week_0 time point.

```{r table-87-shapiro-wilks, include = TRUE, eval = TRUE}
# Data normality assumption

# Data table: Myeloid cells
SubDat[[FigNam]][["Data_Normality_Assumption"]] %>%
  mutate(., across(where(is.numeric), ~ round(.x, 4))) %>%
  mutate(., across("p", ~ ifelse(.x < 0.0001, "<0.0001", format(.x, scientific = FALSE)))) %>%
  group_by(Diet) %>%
  # Build a table
  gt(rowname_col = "Time_point") %>%
    tab_stubhead(label = "Weeks p.i.") %>%
    tab_header(
      title = md("**Appendix Table 87**"),
      subtitle = md("**Univariate Shapito-Wilks test results**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    )
```

```{r qq-plot-figure-S1a, include = TRUE, eval = TRUE, fig.align = "center", fig.height = 9, out.width = "95%", fig.cap = "Fig.S1a-1: QQ-plots of myeloid cell counts split into groups by predictor variables"}
SubDat[[FigNam]][["QQ plot: by groups"]]
```


#### Homogeneity of variance

The assessment of homogeneity of variance was conducted by Levene's test for the dataset split by the within-subject factor ("Time_point"). The analysis output showed that assumption of homogeneity between groups held for each week (Appendix table `r TabNum + 2`).

```{r table-88-homogeneity-of-variance, include = TRUE, eval = TRUE}
# Homogenetiy of variance test: Levene's test outcome
SubDat[[FigNam]][["Homogeneity_of_Variance"]] %>%
  mutate(., across(where(is.numeric), ~ round(.x, 4))) %>%
  mutate(., across("p", ~ ifelse(.x < 0.0001, "<0.0001", format(.x, scientific = FALSE)))) %>%
  # Build a table
  gt(rowname_col = "Time_point") %>%
    tab_stubhead(label = "Weeks p.i.") %>%
    tab_header(
      title = md("**Appendix Table 88**"),
      subtitle = md("**Assessment of homogeneity of variance by week**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    )
```


#### Outliers

We found no outliers in the data (Appendix table `r TabNum + 3`).

```{r table-89-outlier-test, include = TRUE, eval = TRUE}
# Data outliers
if (nrow(SubDat[[FigNam]][["Outliers"]]) == 0) {
  
  data.frame("Outcome" = "No outliers detected") %>%
  # Build a table
    gt() %>%
      tab_header(
        title = md("**Appendix Table 89**"),
        subtitle = md("**List of possible outliers**")
      ) %>%
      tab_style(
        style = cell_text(align = "center"),
        locations = cells_column_labels())
  
} else {
  
  SubDat[[FigNam]][["Outliers"]] %>%
    # Build a table
    gt(rowname_col = "Group") %>%
      tab_stubhead(label = "Groups") %>%
      tab_header(
        title = md("**Appendix Table 89**"),
        subtitle = md("**List of possible outliers**")
      ) %>%
      tab_style(
        style = cell_text(align = "center"),
        locations = cells_column_labels()
      )
}
```

#### Homogeneity of covariance

There was a statistically significant violation of the assumption of homogeneity of covariance as the p-value was <0.0001. We settled for a robust two-way ANOVA to buffer some of the effect of this violation, but only low p-values can be regarded as significant (Appendix table `r TabNum + 4`).

```{r table-90-analysis-of-covariance, include = TRUE, eval = TRUE}
SubDat[[FigNam]][["Box's M test"]] %>%
  dplyr::select(-method) %>%
  mutate(., across(where(is.numeric), ~ round(.x, 4))) %>%
  mutate(., across("p.value", ~ ifelse(.x < 0.0001, "<0.0001", format(.x, scientific = FALSE)))) %>%
  # Build a table
  gt(rowname_col = "Group") %>%
    tab_stubhead(label = "Groups") %>%
    tab_header(
      title = md("**Appendix Table 89**"),
      subtitle = md("**Box's M-test for homogenetiy of covariance**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    )
```

#### Sphericity

We ran a standard two-way ANOVA on the untransformed data to obtain information on the Mauchly's test of sphericity, which showed a clear violation of this assumption. This required the application of Greenhouse-Geisser sphericity correction to the data (Appendix table `r TabNum + 5`).

```{r table-91-mauchly-s-test, , include = TRUE, eval = TRUE}
SubDat[[FigNam]][["Spericity assumption"]][["Mauchly's Test for Sphericity"]] %>% # Greenhouse-Geisser sphericity correction
  mutate(., across(where(is.numeric), ~ round(.x, 4))) %>%
  mutate(., across("p", ~ ifelse(.x < 0.0001, "<0.0001", format(.x, scientific = FALSE)))) %>%
  # Build a table
  gt(rowname_col = "Group") %>%
    tab_stubhead(label = "Groups") %>%
    tab_header(
      title = md("**Appendix Table 89**"),
      subtitle = md("**Mauchly's Test for Sphericity**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    )
```


### Two-way analysis

Based on the assumption analysis, we decided to apply a `r SubDat[[FigNam]][["Assumption_Test_Outcome"]]["2-way_analysis_applied",]` to the dataset to determine the effects of "Diet" and time post infection ("Time_point") on mouse weight over time (Appendix table `r TabNum + 6`). The test output showed that both individual predictors were statistically significant, with the two-way interactions also being statistically significant.

```{r table-91-two-way-anova, include = TRUE, eval = TRUE}
# Two-way ANOVA
SubDat[[FigNam]][["2-way_ANOVA_post_hoc"]] %>%
  capture.output(.) %>%
  trimws(.) %>%
  strsplit(., " ") %>%
  lapply(., function(yy) {
    
    yy[yy != ""]
  }) %>%
  .[c(6:8)] %>%
  lapply(., function(zz) {
    
    zz %>%
      as.data.frame(.) %>%
      t(.) %>%
      as.data.frame(.)
  }) %>%
  bind_rows(.) %>%
  set_colnames(c("Predictor", "value", "p.value")) %>%
  tibble::remove_rownames(.) %>%
  mutate(., across(where(Hmisc::all.is.numeric), as.numeric)) %>%
  # Build a table
  gt(rowname_col = "Predictor") %>%
    tab_stubhead(label = "Predictor") %>%
    tab_header(
      title = md("**Appendix Table 91**"),
      subtitle = md("**Robust two-way ANOVA**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    )
```

For the analysis of the simple main effect for each respective between-subject factor, we performed `r SubDat[[FigNam]][["Assumption_Test_Outcome"]]["1-way_analysis_applied",]` with the between-subject factor ("Diet") split by the within-subject factor ("Time_point"). The results showed that "Diet" caused statistically significant differences with the exception of "Week_0" (Table `r TabNum + 7`).

```{r table-92-one-way-anova, include = TRUE, eval = TRUE}
# One-way ANOVA
# Data table
SubDat[[FigNam]][["1-way_ANOVA_simple_simple_main_effect"]] %>%
  mutate(., across(where(is.numeric), ~ round(.x, 4))) %>%
  mutate(., across(c("p.value"), ~ ifelse(.x < 0.0001, "<0.0001", format(.x, scientific = FALSE)))) %>%
  # Build a table
  gt(rowname_col = "Time_point") %>%
    tab_stubhead(label = "Weeks p.i.") %>%
    tab_header(
      title = md("**Appendix Table 92**"),
      subtitle = md("**Robust one-way ANOVA**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    )
```

For the pairwise comparison, we applied a `r SubDat[[FigNam]][["Assumption_Test_Outcome"]]["Pairwise_analysis_applied",]` (Table `r TabNum + 8`). Since the "Diet" predictor only had two factor levels, the output showed the same result as the `r SubDat[[FigNam]][["Assumption_Test_Outcome"]]["1-way_analysis_applied",]` above (Table `r TabNum + 7`).

```{r table-94-pairwise-comparison, include = TRUE, eval = TRUE}
# Pairwise comparison
# This table is only for Diet as predictor and only rows with significant p-values
SubDat[[FigNam]][["Pairwise_simple_simple_comparison"]] %>%
  mutate(., across(where(is.numeric), ~ round(.x, 4))) %>%
  # .[.[["p.value"]] <0.05,] %>%
  # .[.$Predictor == "Route",] %>%
  mutate(., across(where(is.numeric), ~ round(.x, 4))) %>%
  mutate(., across(c("p.value"), ~ ifelse(.x < 0.0001, "<0.0001", format(.x, scientific = FALSE)))) %>%
  # Build a table
  gt(rowname_col = "Time_point") %>%
    tab_stubhead(label = "Weeks p.i.") %>%
    tab_header(
      title = md("**Appendix Table 93**"),
      subtitle = md("**Robust one-way ANOVA - sig. summary**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    ) 
```


## Conclusion

In conclusion, "Diet" was a potent predictor for mouse weight gain over time.

```{r update-table-number-figure-S1a}
TabNum <- TabNum + 8
```
