---
title: "Figure S1c"
author: "Johannes S. P. Doehl"
date: "2024-06-21"
output: pdf_document
---

~~~{=comment}
# Setup
~~~
```{r setup-figure-S1c, echo = FALSE, include = FALSE, eval = TRUE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE, eval = TRUE, cache = TRUE, dev = "png", warning = FALSE, message = FALSE, fig.pos = "H", out.extra = "")
options(java.parameters = "-Xmx200000m")
Sys.setenv(JAVA_HOME='C:\\Program Files\\Java\\jre-1.8') # Needed to get around loading error of rJava
```

```{r extract-relevant-data-for-figure-S1c}
# Figure name
FigNam <- "Figure S1c"

# Total mice used
TotN <- DatCntSpt %>%
  .[names(.) %like% FigNam] %>%
  .[[1]] %>%
  nrow(.)

# Total N per group
GrpN <- table(DatCntSpt %>%
        .[names(.) %like% FigNam] %>%
        .[[1]] %>%
        .[, "Diet"]) %>%
  paste(gsub("_.*", "", names(.)),
        ., 
        sep = "=", 
        collapse = ", ")

# Dispersion ratio
DispRat <- RegModAIC %>%
  .[names(.) %like% FigNam] %>%
  .[[1]] %>%
  .[[1]] %>%
  .[["Dispersion test 2"]] %>%
  c(.) %>%
  lapply(., function(xx) {
    round(xx, 4)
    }) 

# Pseudo-R2
pR2 <- RegModAIC %>%
  .[names(.) %like% FigNam] %>%
  .[[1]] %>%
  .[[1]] %>%
  .[["Pseudo R-square"]] %>%
  .[["Pseudo.R.squared.for.model.vs.null"]] %>%
  as.data.frame(.) %>%
  tibble::rownames_to_column(.) %>%
  .[3,]
```

We analyzed a total of *N*=`r TotN` BALB/c mice (`r GrpN`). These were the same mice as analyzed in figure 2a for parasite dissemination events. Here, we quantified sand fly feeding success by counting the number of visibly fed sand flies. For the data analysis we tested several Poisson and negative binomial-type regression models. Based on the Akaike information criterion (AIC) we selected a `r RegModAIC %>% .[names(.) %like% FigNam] %>% .[[1]] %>% names(.) %>% gsub("*- ", "", .)` regression model for the data analysis post data re-scaling. The model fitted the data well producing no statistically significant departure from 1 for its dispersion ratio (`r paste(names(DispRat), DispRat, sep = ": ", collapse = ", ")`), but showing only a small pseudo-R^2^ (`r paste(pR2, collapse = ": ")`). The model output showed that "Diet" was not a statistically significant predictor, suggesting that the nutritional state of the mice did not impact the sand flies ability to feed on them (Appendix table `r TabNum + 1`).

```{r table-168-regression-analysis-of-counts, include = TRUE, eval = TRUE}
# Model output
RegModAIC %>%
  .[names(.) %like% FigNam] %>%
  .[[1]] %>% # pulls out the data from the list
  .[[1]] %>% # picks the only model output left
  .[["Regression model"]] %>%
  summary(.) %>%
  stats::coef(.) %>%
  as.data.frame(.) %>%
  mutate(., "sig." = .[["Pr(>|z|)"]] %>%
           symnum(.,
                  cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 0.1, 1),
                  symbols = c("****", "***", "**", "*", "+", "ns"))) %>%
  mutate(., across(where(is.numeric), ~ round(.x, 4))) %>%
  mutate(., across("Pr(>|z|)", ~ ifelse(.x < 0.0001, "<0.0001", format(.x, scientific = FALSE)))) %>%
  # Build a table
  gt(rownames_to_stub = TRUE) %>%
    tab_stubhead(label = "Predictors") %>%
    tab_header(
      title = md("**Appendix Table 168**"),
      subtitle = md("**Negative binomial regression model output**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    ) %>%
    cols_align(
      align = "right",
      columns = everything()
    )
```

The pairwise comparison based on the estimated marginal means showed that sand flies fed just as well on well-nourished animals as on malnourished animals excluding any unintended parasite inoculation bias due to the nutritional state (Appendix table `r TabNum + 2` and `r TabNum + 3`).

```{r table-169-pairwise-comparison, include = TRUE, eval = TRUE}
RegModAIC %>%
  .[names(.) %like% FigNam] %>%
  .[[1]] %>% # pulls out the data from the list
  .[[1]] %>% # picks the only model output left
  .[["Pairwise Comparison of EMMs"]] %>%
  as.data.frame(.) %>%
  mutate(., "sig." = .[["p.value"]] %>%
           symnum(.,
                  cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 0.1, 1),
                  symbols = c("****", "***", "**", "*", "+", "ns"))) %>%
  mutate(., across(where(is.numeric), ~ round(.x, 4))) %>%
  mutate(., across("p.value", ~ ifelse(.x < 0.0001, "<0.0001", format(.x, scientific = FALSE)))) %>%
  # Build a table
  gt(rowname_col = "contrast") %>%
    tab_stubhead(label = "Predictor pairs") %>%
    tab_header(
      title = md("**Appendix Table 169**"),
      subtitle = md("**Pairwise comparison based on estimated marginal means**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    ) %>%
    cols_align(
      align = "right",
      columns = everything()
    )
```

```{r table-170-pairwise-comparison-letter, include = TRUE, eval = TRUE}
RegModAIC%>%
  .[names(.) %like% FigNam] %>%
  .[[1]] %>% # pulls out the data from the list
  .[[1]] %>% # picks the only model output left
  .[["Pairwise Comparison of EMMs (letters)"]] %>%
  mutate(., across(where(is.numeric), ~ round(.x, 4))) %>%
  # Build a table
  gt() %>%
    tab_header(
      title = md("**Appendix Table 170**"),
      subtitle = md("**Pairwise comparison letter code**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    ) %>%
    cols_align(
      align = "right",
      columns = everything()
    )
```


```{r update-table-number-figure-S1c}
TabNum <- TabNum + 3
```
