---
title: "Figure 4d"
author: "Johannes S. P. Doehl"
date: "2024-07-10"
output: pdf_document
---

~~~{=comment}
# Setup
~~~
```{r setup-figure-4d, echo = FALSE, include = FALSE, eval = TRUE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE, eval = TRUE, cache = TRUE, dev = "png", warning = FALSE, message = FALSE, fig.pos = "H", out.extra = "")
options(java.parameters = "-Xmx200000m")
Sys.setenv(JAVA_HOME='C:\\Program Files\\Java\\jre-1.8') # Needed to get around loading error of rJava
```

```{r extract-relevant-data-for-figure-4d}
# Figure name
FigNam <- "Figure 4d"

# Name of tissue processed
TisNam <- Res2w %>%
  .[names(.) %like% FigNam] %>%
  names(.) %>%
  gsub(".*- ", "", .)

# Subset data for figure 4d
SubDat <- Res2w %>%
  .[names(.) %like% FigNam] %>%
  purrr::set_names(TisNam)

# Total mice used
TotN <- lapply(Dat2wSpt %>%
                 .[names(.) %like% FigNam], function(zz) {
  nrow(zz)
}) %>%
  purrr::set_names(TisNam)

# Total N per group
GrpN <- mapply(function(xx, yy) {
  table(yy$Diet, yy$Route) %>%
    as.data.frame(.) %>%
    unite(., c(Var1, Var2), col = "Group", sep = "_") %>%
    tibble::column_to_rownames("Group") %>%
    set_colnames(xx)
  
}, xx = TisNam, yy = Dat2wSpt %>%
                 .[names(.) %like% FigNam], SIMPLIFY = FALSE) %>%
  bind_cols(.)
```


## Data analysis

Here, we present statistical comparison of serum cytokine concentrations (in pg/ml) for a total of *N*=`r unique(TotN)` well-nourished (WN) or malnourished (MN) BALB/c mice infected by sand fly bite (SF) or not (Ctrl) (Appendix table `r TabNum + 1`). A total of `r length(TisNam)` targets were measured (`r stringi::stri_replace_last(paste(TisNam, collapse = ", "), fixed = ",", " &")`) per mouse.

```{r table-148-sample-size-information, include = TRUE, eval = TRUE}
# Sample size information by cell type
rbind(TotN %>%
        bind_cols(.) %>%
        as.data.frame(.) %>%
        set_rownames("Total N"),
      GrpN
      ) %>%
  t(.) %>%
  as.data.frame.matrix(.) %>%
  cbind(.,
        sapply(TisNam, function(zz) {
          SubDat[[zz]][["Assumption_Test_Outcome"]]["Dat_Transformation", "Outcome"]
          }) %>%
          as.data.frame(.) %>%
          rename("Transformation" = ".")
        ) %>%
  # Build a table
  gt(rownames_to_stub = TRUE) %>%
    tab_stubhead(label = "Targets") %>%
    tab_header(
      title = md("**Appendix Table 148**"),
      subtitle = md("**Summary information**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    ) %>%
    cols_align(
      align = "right",
      columns = everything()
    )
```

We needed to analyze the data with a two-way approach to account for the two predictors, "Diet" and "Route", both of which were between-subject factors.

For a two-way ANOVA, we had to assess the data for compliance with assumptions:

-   Data normality
-   Homogeneity of variance
-   No significant outliers

## Assumption analyses

Please, note that all assumption test results shown were post data transformation, where applicable (Appendix table `r TabNum + 1`).

### Data normality

The assessment of the untransformed data distribution for each group was conducted by Shapiro-Wilks test and QQ-plot after splitting the data by both predictors. Note that all groups consisted of \<30 data points (range[7 to 13]) (Appendix table `r TabNum + 1`), which made groups too small to assess data distribution reliably by Shapiro-Wilks test. We executed the test anyway as an indicator of gross departure of data normality. Thus, we performed the analyses by Shapiro-Wilks test (Appendix table `r TabNum + 2`) and QQ-plots (Fig.S1a-1). We found occasional departure of normality, which could not all be remedied by data transformation.

```{r table-149-shapiro-wilks, include = TRUE, eval = TRUE}
# Data normality assumption
lapply(TisNam, function(qq) {
  SubDat[[qq]][["Data_Normality_Assumption"]]
  }) %>%
  purrr::set_names(TisNam) %>%
  bind_rows(., .id = "Targets") %>%
  mutate(., across(where(is.numeric), ~ round(.x, 4))) %>%
  mutate(., across("p", ~ ifelse(.x < 0.0001, "<0.0001", format(.x, scientific = FALSE)))) %>%
  group_by(Targets) %>%
  # Build a table
  gt() %>%
    tab_header(
      title = md("**Appendix Table 149**"),
      subtitle = md("**Univariate Shapito-Wilks test results**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    )
```

```{r qq-plot-figure-4d, include = TRUE, eval = TRUE, fig.align = "center", fig.height = 9, out.width = "95%", fig.cap = "Fig.4d-1: QQ-plots of serum cytokine data: A) IFN-gamma, B) IL-10, C) IL-12p70, D) IL-17A, E) IL-18, F) IL-2, G) IL-22, H) IL-4, I) IL-5, J) IL-6 and K) TNF-alpha"}
ggarrange(SubDat[["IFN-gamma"]][["QQ plot: by groups"]],
          SubDat[["IL-10"]][["QQ plot: by groups"]],
          SubDat[["IL-12p70"]][["QQ plot: by groups"]],
          SubDat[["IL-17A"]][["QQ plot: by groups"]],
          SubDat[["IL-18"]][["QQ plot: by groups"]],
          SubDat[["IL-2"]][["QQ plot: by groups"]],
          SubDat[["IL-22"]][["QQ plot: by groups"]],
          SubDat[["IL-4"]][["QQ plot: by groups"]],
          SubDat[["IL-5"]][["QQ plot: by groups"]],
          SubDat[["IL-6"]][["QQ plot: by groups"]],
          SubDat[["TNF-alpha"]][["QQ plot: by groups"]],
          labels = LETTERS[1:11],
          ncol = 3, 
          nrow = 4)
```


### Homogeneity of variance

The assessment of homogeneity of variance was conducted by Levene's test for the dataset. The analysis output showed that assumption of homogeneity between groups held for all cytokines (Appendix table `r TabNum + 3`).

```{r table-150-homogeneity-of-variance, include = TRUE, eval = TRUE}
# Homogeneity of variance test: Levene's test outcome
lapply(TisNam, function(qq) {
  SubDat[[qq]][["Homogeneity_of_Variance"]]
  }) %>%
  purrr::set_names(TisNam) %>%
  bind_rows(., .id = "Targets") %>%
  mutate(., "sig." = .[["p"]] %>%
           symnum(.,
                  cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 0.1, 1),
                  symbols = c("****", "***", "**", "*", "+", "ns"))) %>%
  mutate(., across(where(is.numeric), ~ round(.x, 4))) %>%
  mutate(., across("p", ~ ifelse(.x < 0.0001, "<0.0001", format(.x, scientific = FALSE)))) %>%
  # Build a table
  gt(rowname_col = "Targets") %>%
    tab_stubhead(label = "Targets") %>%
    tab_header(
      title = md("**Appendix Table 150**"),
      subtitle = md("**Assessment of homogeneity of variance by target**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    )
```


### Outliers

Conversely, we detected outliers for 9 of the 11 cytokines (Appendix table `r TabNum + 4`). Only two of these outliers were classified as extreme (one each for IL-12p70 and IL-6).

```{r table-151-outlier-test, include = TRUE, eval = TRUE}
# Data outliers
lapply(TisNam, function(qq) {
    SubDat[[qq]][["Outliers"]]
  }) %>%
  purrr::set_names(TisNam) %>%
  bind_rows(., .id = "Targets") %>%
  group_by(Targets) %>%
  # Build a table
  gt(rowname_col = "Group") %>%
    tab_stubhead(label = "Groups") %>%
    tab_header(
      title = md("**Appendix Table 151**"),
      subtitle = md("**List of possible outliers**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    )
```

## Two-way analysis

Based on the assumption analysis, we decided to applied an appropriate two-way tests to the respective datasets (Appendix table `r TabNum + 5`).

```{r table-152-applied-tests, include = TRUE, eval = TRUE}
# Applied tests
cbind(
  lapply(SubDat, function(aa) {
    aa[["Assumption_Test_Outcome"]]["2-way_ANOVA_Test_applied", "Outcome"]
  }) %>%
    bind_cols(.) %>%
    t(.) %>%
    set_colnames("Two-way test"),
  lapply(SubDat, function(aa) {
    aa[["Assumption_Test_Outcome"]]["1-way_ANOVA_Test_applied", "Outcome"]
  }) %>%
    bind_cols(.) %>%
    t(.) %>%
    set_colnames("One-way test"),
  lapply(SubDat, function(aa) {
    aa[["Assumption_Test_Outcome"]]["Pairwise_comparison_applied", "Outcome"]
  }) %>%
    bind_cols(.) %>%
    t(.) %>%
    set_colnames("Pairwise")
  ) %>%
  as.data.frame(.) %>%
   # Build a table
  gt(rownames_to_stub = TRUE) %>%
    tab_stubhead(label = "Targets") %>%
    tab_header(
      title = md("**Appendix Table 152**"),
      subtitle = md("**Applied tests**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    )
```

The respective two-way analyses showed that there was no statistical significant differences detected for IL-22 (Appendix table `r TabNum + 6`), suggesting that this cytokine was not statistically significantly affected by either nutritional status and/or infection route. All other targets had detectable statistical significant differences mostly in relation to "Diet" with the exception of IFN-$\gamma$, where only the interaction term of "Diet" and "Route" was statistically significant. IL-10 and IL-17A also showed statistical significance for the infection route, but not for the interaction term. Conversely, IL-6 showed a statistically significant interaction between "Diet" and "Route", while not for "Route" on its own. Finally, TNF-$\alpha$ showed statistical significance for both predictors as much as their interaction. This suggested that most serum cytokine levels were affected by the nutritional state of the mice, while some changed their serum concentration with respect to the infection.

```{r table-153-two-way-anova, include = TRUE, eval = TRUE}
# Two-way ANOVA
lapply(TisNam, function(qq) {
  
  if (SubDat[[qq]][["Assumption_Test_Outcome"]]["2-way_ANOVA_Test_applied", "Outcome"] %like% "Robust") {
    
    SubDat[[qq]][["2-way_ANOVA_Type"]] %>%
      capture.output(.) %>%
      trimws(.) %>%
      strsplit(., " ") %>%
      lapply(., function(yy) {
        
        yy[yy != ""]
      }) %>%
      .[c(5:7)] %>%
      lapply(., function(zz) {
        
        zz %>%
          as.data.frame(.) %>%
          t(.) %>%
          as.data.frame(.)
      }) %>%
      bind_rows(.) %>%
      set_colnames(c("Effect", "value", "p.value")) %>%
      tibble::remove_rownames(.) %>%
      mutate(., across(where(Hmisc::all.is.numeric), as.numeric)) %>%
      mutate(., DFn = c(1,1,1),
                DFd = c(31,31,31),
                .before = "value") %>%
      mutate(., "p<.05" = .[["p.value"]] %>%
           symnum(.,
                  cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 0.1, 1),
                  symbols = c("****", "***", "**", "*", "+", " "))) %>%
      mutate(., "ges" = c(NA,NA,NA)) %>%
      rename("Statistic" = "value")
    
  } else if (SubDat[[qq]][["Assumption_Test_Outcome"]]["2-way_ANOVA_Test_applied", "Outcome"] %like% "regression") {
    
    SubDat[[qq]][["2-way_ANOVA_Type"]] %>%
      as.data.frame(.) %>%
      .[1:3,] %>%
      tibble::rownames_to_column(., var = "Effect") %>%
      dplyr::select(-c("Sum Sq", "Mean Sq")) %>%
      mutate(., DFd = c(31,31,31),
                .after = "Df") %>%
      mutate(., "p<.05" = .[["Pr(>F)"]] %>%
           symnum(.,
                  cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 0.1, 1),
                  symbols = c("****", "***", "**", "*", "+", " "))) %>%
      mutate(., "ges" = c(NA,NA,NA)) %>%
      rename("DFn" = "Df",
             "Statistic" = "F value",
             "p.value" = "Pr(>F)") %>%
      mutate(., across(where(Hmisc::all.is.numeric), as.numeric))
    
  } else {
    
    SubDat[[qq]][["2-way_ANOVA_Type"]] %>%
      as.matrix(.) %>%
      as.data.frame(.) %>%
      rename("Statistic" = "F",
             "p.value" = "p") %>%
      mutate(., across(where(Hmisc::all.is.numeric), as.numeric)) %>% 
      mutate(., across("p<.05", noquote))
  }
  }) %>%
  purrr::set_names(TisNam) %>%
  bind_rows(., .id = "Targets") %>%
  tibble::remove_rownames(.) %>%
  mutate(., across(where(is.numeric), ~ round(.x, 4))) %>%
  mutate(., across("p.value", ~ ifelse(.x < 0.0001, "<0.0001", format(.x, scientific = FALSE)))) %>%
  group_by(Targets) %>%
  # Build a table
  gt(rowname_col = "Predictor") %>%
    tab_stubhead(label = "Predictor") %>%
    tab_header(
      title = md("**Appendix Table 153**"),
      subtitle = md("**Two-way analysis**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    )
```

For the analysis of the simple main effect for each respective between-subject factor for each target, we performed one-way analyses with the data split by always by the predictor that was not used as within the function. The results showed that statistical significance for each cytokine, including IL-22, which had not shown statistical significance in the two-way analysis (Table `r TabNum + 7`). Results were quite varied between cytokines, frequently showing statistically significant differences between well-nourished uninfected and sand fly infected mice (IFN-$\gamma$, IL-10, IL-17A, IL-22 & IL-5) and well- and malnourished uninfected control groups showed statistical significant differences for all measured serum cytokines. The latter indicated that the nutritional status of the mice altered their steady state levels of measured cytokines, increasing them significantly in the malnourished group; a difference negated after parasite infection by sand fly.

```{r table-154-one-way-anova, include = TRUE, eval = TRUE}
# One-way ANOVA
# Data table
lapply(TisNam, function(qq) {
  
  if (SubDat[[qq]][["Assumption_Test_Outcome"]]["1-way_ANOVA_Test_applied", "Outcome"] %like% "Robust") {
    
    lapply(SubDat[[qq]][["1-way_ANOVA_simple_simple_main_effect"]], function(rr) {
      
      rr %>%
        as.data.frame(.) %>%
        rename(any_of(c("Factor" = "Diet",
                      "Factor" = "Route")))
        
    }) %>%
      bind_rows(., .id = "Effect") %>%
      mutate(., across("Effect", ~ gsub(".*: ", "", .x))) %>%
      dplyr::select(-starts_with("effsize_ci")) %>%
      .[, c(1,2,3,4,5,6,8,7)] %>%
      rename(c("statistic" = "test"))
    
  } else if (SubDat[[qq]][["Assumption_Test_Outcome"]]["1-way_ANOVA_Test_applied", "Outcome"] %like% "Standard") {
    
    lapply(SubDat[[qq]][["1-way_ANOVA_simple_simple_main_effect"]], function(rr) {
      
      rr %>%
        as.data.frame(.) %>%
        rename(any_of(c("Factor" = "Diet",
                      "Factor" = "Route")))
        
    }) %>%
      bind_rows(., .id = "Effect") %>%
      mutate(., across("Effect", ~ gsub(".*: ", "", .x))) %>%
      .[, c(1,2,5,3,4,6,7,8)] %>%
      rename(c("statistic" = "F",
               "df1" = "DFn",
               "df2" = "DFd",
               "p.value" = "p",
               "effsize" = "ges",
               "Sig." = "p<.05")) %>% 
      mutate(., across("Sig.", noquote))
  }
  }) %>%
  purrr::set_names(TisNam) %>%
  bind_rows(., .id = "Targets") %>%
  mutate(., across(where(is.numeric), ~ round(.x, 4))) %>%
  mutate(., across(c("p.value"), ~ ifelse(.x < 0.0001, "<0.0001", format(.x, scientific = FALSE)))) %>%
  mutate(., across("Sig.", ~ ifelse(.x == "", "ns", .x))) %>%
  mutate(., across("Sig.", noquote)) %>%
  group_by(Targets) %>%
  # Build a table
  gt(rowname_col = "Factor") %>%
    tab_stubhead(label = "Factor") %>%
    tab_header(
      title = md("**Appendix Table 154**"),
      subtitle = md("**Simple main effect analysis**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    )
```

Since both predictors had only two levels each, the pairwise comparison reflected the observations of the one-way analyses above (Table `r TabNum + 8`).

```{r table-155-pairwise-comparison, include = TRUE, eval = TRUE}
# Pairwise comparison
# This table is only for Diet as predictor and only rows with significant p-values

lapply(TisNam, function(pp) {
  
  if (SubDat[[pp]][["Assumption_Test_Outcome"]]["Pairwise_comparison_applied", "Outcome"] %like% "Linear") {
    
    mapply(function(qq, rr) {
      qq %>%
        rename(any_of(c("Factor" = "Diet",
                        "Factor" = "Route"))) %>%
        { if (rr %like% "Route") {
            mutate(., 
                 "contrast" = paste(paste(.$Factor, "Ctrl"),
                                    paste(.$Factor, "SF"),
                                    sep = " - "),
                 .before = "Factor")
          } else {
            mutate(., 
                 "contrast" = paste(paste("WN", .$Factor),
                                    paste("MN", .$Factor),
                                    sep = " - "),
                 .before = "Factor")
          }
        }
    }, qq = SubDat[[pp]][["Pairwise_simple_simple_comparison"]],  
       rr = names(SubDat[[pp]][["Pairwise_simple_simple_comparison"]]), 
       SIMPLIFY = FALSE) %>%
      bind_rows(.) %>%
      mutate(., 
             "df" = rep(31, times = 4),
             .after = "contrast") %>%
      rename("statistic" = "psihat") %>%
      .[, c(1,2,4,7,8)] %>%
      arrange(contrast)
    
  } else if (SubDat[[pp]][["Assumption_Test_Outcome"]]["2-way_ANOVA_Test_applied", "Outcome"] %like% "regression") {
    
    SubDat[[pp]][["Pairwise_simple_simple_comparison"]][["Pairwise comparison"]] %>% 
      as.data.frame(.) %>% 
      mutate(., "Sig." = .[["p.value"]] %>%
           symnum(.,
                  cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 0.1, 1),
                  symbols = c("****", "***", "**", "*", "+", "ns"))) %>%
      dplyr::filter(!contrast %in% "WN Ctrl - MN SF", # Remove the pairs that are of no interest
                    !contrast %in% "MN Ctrl - WN SF") %>%
      .[, c(1,4,2,6,7)] %>%
      rename("statistic" = "estimate") %>%
      arrange(contrast)
    
  } else {
    
    mapply(function(qq, rr) {
      qq %>%
        rename(any_of(c("Factor" = "Diet",
                        "Factor" = "Route"))) %>%
        { if (rr %like% "Route") {
            mutate(., 
                 "contrast" = paste(paste(.$Factor, .$group1),
                                    paste(.$Factor, .$group2),
                                    sep = " - "),
                 .before = "Factor")
          } else {
            mutate(., 
                 "contrast" = paste(paste(.$group1, .$Factor),
                                    paste(.$group2, .$Factor),
                                    sep = " - "),
                 .before = "Factor")
          }
        }
    }, qq = SubDat[[pp]][["Pairwise_simple_simple_comparison"]],  
       rr = names(SubDat[[pp]][["Pairwise_simple_simple_comparison"]]), 
       SIMPLIFY = FALSE) %>%
      
      bind_rows(.) %>%
      dplyr::select(-c("Factor",
                       "term",
                       ".y.",
                       "group1",
                       "group2",
                       "p")) %>%
      rename("p.value" = "p.adj",
             "Sig." = "p.adj.signif") %>%
      as.data.frame(.) %>%
      arrange(contrast) %>%
      mutate(., across("Sig.", noquote))
    
    
  }
  }) %>%
  purrr::set_names(TisNam) %>%
  bind_rows(., .id = "Target") %>%
  mutate(., across(where(is.numeric), ~ round(.x, 4))) %>%
  mutate(., across(c("p.value"), ~ ifelse(.x < 0.0001, "<0.0001", format(.x, scientific = FALSE)))) %>%
  group_by(Target) %>%
  # Build a table
  gt(rowname_col = "contrast") %>%
    tab_stubhead(label = "contrast") %>%
    tab_header(
      title = md("**Appendix Table 155**"),
      subtitle = md("**Pairwise comparison**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    ) 
```


## Statistical power

Considering the small group sizes (Appendix tables `r TabNum + 1`), we wanted to ensure that the study design was not significantly statistically underpowered. Thus, we performed a retrospective sample size and power analysis on the data by target.

```{r subsetting-the-power-calculation-data-for-figure-4d}
# Subset data for figure 4d
SubDat <- PwrCal2w %>%
  .[names(.) %like% FigNam] %>%
  purrr::set_names(TisNam)
```


### Effect size estimation based on partial eta^2^

Effect sizes were calculated by predictor and the different potential interaction combinations of them. Appendix tables `r TabNum + 9` shows the respective effect sizes. Note that upper ends of the confidence intervals were automatically set to 1 for this type of calculation. Large effect sizes reflected statistically meaningful differences in the data analysis. The partial eta^2^ values from the effect size calculation were then used for the retrospective power calculations.

```{r table-156-effect-size, include = TRUE, eval = TRUE}
#Effect size
lapply(SubDat, function(qq) {
  qq[["Effect_Size"]] %>%
    as.data.frame(.) %>%
    dplyr::select(!CI) 
}) %>%
  bind_rows(., .id = "Target") %>%
  mutate(., across(where(Hmisc::all.is.numeric), as.numeric)) %>%
  mutate(., across(where(is.numeric), ~ format(round(.x, 4), scientific = FALSE))) %>%
  mutate(., "Effect Size" = ifelse(.[["Eta2_partial"]] < 0.01, "very small",
                                           ifelse(.[["Eta2_partial"]] < 0.06, "small",
                                                  ifelse(.[["Eta2_partial"]] < 0.14, "medium", "large")))) %>%
  group_by(Target) %>%
  # Build a table
  gt(rowname_col = "Parameters") %>%
    tab_stubhead(label = "Predictors") %>%
    tab_header(
      title = md("**Appendix Table 156**"),
      subtitle = md("**Effect size**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    )
```


### Retrospective minimum total sample size estimation for 80% power

The accepted rule of thumb is to have at least 80% (0.8 as ratio) statistical power in once data. For small mean differences within data of a high level of complexity, this is often hard to achieve, because of cost and ability to manage large sample sizes. In most instances, proposed sample sizes based on our data, suggested that actual sample sizes were frequently too small for a chance of detecting statistical significant differences, particularly, for the interaction of predictors (Appendix tables `r TabNum + 10`).

```{r table-157-sample-size, include = TRUE, eval = TRUE}
# Retrospective sample size calculation
lapply(SubDat, function(qq) {
  qq[["Sample_Size_for_80%_Power"]] %>%
    bind_rows(., .id = "Parameters")
}) %>%
  purrr::set_names(TisNam) %>%
  bind_rows(., .id = "Target") %>%
  mutate(., across(where(Hmisc::all.is.numeric), as.numeric)) %>%
  mutate(., across(where(is.numeric), ~ format(round(.x, 2), scientific = FALSE))) %>%
  group_by(Target, Parameters) %>%
  # Build a table
  gt(rowname_col = "effect") %>%
    tab_stubhead(label = "Effect") %>%
    tab_header(
      title = md("**Appendix Table 157**"),
      subtitle = md("**Minimum optimal sample size calculation**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    ) %>%
    tab_style(
      style = cell_text(weight = "bold"),
      location = cells_body(
        columns = "n.total")
    )
```


### Retrospective calculation of statistical power in our data analysis

With the exception of urea nitrogen, we generally observed that the our study was underpowered for the detection of a statistical significant difference, in particular, for the infection route and the interaction with "Diet" (Appendix Table `r TabNum + 11`). Even so, excessively large proposed sample sizes and small retrospective statistical power can also be an indicator that there was in fact no meaningful biological difference to be found. Considering that statistical significant difference were still found so frequently, suggested that the above observations were real.

```{r table-158-power-calculation, include = TRUE, eval = TRUE}
# Retrospective power calculation
lapply(SubDat, function(qq) {
  qq[["Statistical_Power_of_Data"]] %>%
    bind_rows(., .id = "Parameters")
}) %>%
  purrr::set_names(TisNam) %>%
  bind_rows(., .id = "Target") %>%
  mutate(., across(where(Hmisc::all.is.numeric), as.numeric)) %>%
  mutate(., across(where(is.numeric), ~ format(round(.x, 2), scientific = FALSE))) %>%
  group_by(Target, Parameters) %>%
  # Build a table
  gt(rowname_col = "effect") %>%
    tab_stubhead(label = "Effect") %>%
    tab_header(
      title = md("**Appendix Table 158*"),
      subtitle = md("**Statistical power of data**")
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels()
    ) %>%
    tab_style(
      style = cell_text(weight = "bold"),
      location = cells_body(
        columns = "power")
    )
```


```{r update-table-number-figure-4d}
TabNum <- TabNum + 11
```
